A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-05-30 sections/Overview.tex
"worklist judgments"

This section starts by introducing DK's declarative type system. Then
it discusses several techniques that have been used in algorithmic
formulations, and which have influenced our own algorithmic design.
Finally we introduce the novelties of our new algorithm.
In particular the support for inference judgments in
judgment worklists, and a new form of judgment worklist
that the unifies \emph{ordered contexts} and the worklists themselves.

@@

% %-------------------------------------------------------------------------------
% \subsection{Small-Step Unification}
% % Mention some algorithms for unification for 
% % dependent types that use a small-step approach. Credit them later for some ideas that 
% % we also employ.
% There are literals that make use of small-step unification for dependently typed
% inference and reconstruction algorithms~\cite{Reed2009,Abel2011higher}.
% These approaches collect a list of constraints and process one at a time.
% Similar to DK's algorithm, unification variables are used to represent the types to guess.
% In order to solve unification variables to correctly scoped types
% (typically constrained by the context when the variable is introduced),
% the context information should be kept with the variables.
% Judgments that represent unifications of terms might also require such a context
% to keep the constraint collection well-formed.
% 
% As an example, Abel et al. defines their constraint $K$ by
% \begin{gather*}
% \begin{aligned}
% K &::= {\top} \mid {\bot} &&\text{Trivial constraint and inconsistency.}\\
%     & \ \mid \  \Psi\vdash M = N : C &&\text{Unify term $M$ with $N$.}\\
%     & \ \mid \  \Psi\mid R:A \vdash E = E' &&\text{Unify evaluation context $E$ with $E'$.}\\
%     & \ \mid \  \Psi\vdash u\leftarrow M: C &&\text{Solution for $u$ (metavariable) found.}
% \end{aligned}
% \end{gather*}
% where the unification constraints and metavariable solutions are bound by a proper context.
% A unification problem in their system is described by $\Delta \Vdash \overline{K}$,
% where $\Delta$ is a collection of metavariables with their defined scopes.
% This approach clearly states the scroping of each unification problem,
% therefore rules out invalid instantiations to metavariables.
% 
% However, such ``duplicated contexts'' are not ideal for our formalization.
% Following DK's algorithm, an existential variable $\al$ defined in the context
% could be decomposed into a function type $\al[1] \to \al[2]$,
% so the declaration of $\al$ should be replaced by two declarations $\al[1], \al[2]$.
% Such operation causes all the context that contains $\al$ to change,
% which brings difficulty on the synchronization of multiple contexts,
% as variable declarations are not centralized.
% 
% We find a way of encoding multiple contexts by ``compressing'' them into a single worklist.
% As our type inference rules applied to a judgment,
% we typically keep the context or add variable declarations to the old context
% before analysing its sub-judgment(s).
% This enables us to write the judgment worklist in such a form:
% $$\Gm, \{\text{variable declarations}\}, \{\text{judgment}\},
% \{\text{variable declarations}\}, \{\text{judgment}\}, \ldots .$$
% Typically, we simplify the right-most judgment and push back smaller tasks,
% where the old ``context'' automatically gets inherited.
% When some rules solve or partially solve an existential variable,
% we could easily propagate the solution to all the judgments,
% and safely modify its declaration, such as remove from the worklist.
% New judgment(s) may also introduce new \emph{local} variables by declaring
% that variable right before the new judgment(s).
% After the new judgments are satisfied, these local variables are properly recycled.
% 
% % \jimmy{TODO explain the duplicated context with some examples (their judgment form)}
% % Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
% % the formalization harder since requires ``synchronizing'' things ...
% 
%-------------------------------------------------------------------------------
\subsection{Single-Context Worklist Algorithm for Subtyping}

>>>>

This section starts by introducing DK's declarative type system. Then
it discusses several techniques that have been used in algorithmic
formulations, and which have influenced our own algorithmic design.
Finally we introduce the novelties of our new algorithm.
In particular the support for inference judgments in
worklist judgments, and a new form of worklist judgment
that the unifies \emph{ordered contexts} and the worklists themselves.

@@

% %-------------------------------------------------------------------------------
% \subsection{Small-Step Unification}
% % Mention some algorithms for unification for 
% % dependent types that use a small-step approach. Credit them later for some ideas that 
% % we also employ.
% There are literals that make use of small-step unification for dependently typed
% inference and reconstruction algorithms~\cite{Reed2009,Abel2011higher}.
% These approaches collect a list of constraints and process one at a time.
% Similar to DK's algorithm, unification variables are used to represent the types to guess.
% In order to solve unification variables to correctly scoped types
% (typically constrained by the context when the variable is introduced),
% the context information should be kept with the variables.
% Judgments that represent unifications of terms might also require such a context
% to keep the constraint collection well-formed.
% 
% As an example, Abel et al. defines their constraint $K$ by
% \begin{gather*}
% \begin{aligned}
% K &::= {\top} \mid {\bot} &&\text{Trivial constraint and inconsistency.}\\
%     & \ \mid \  \Psi\vdash M = N : C &&\text{Unify term $M$ with $N$.}\\
%     & \ \mid \  \Psi\mid R:A \vdash E = E' &&\text{Unify evaluation context $E$ with $E'$.}\\
%     & \ \mid \  \Psi\vdash u\leftarrow M: C &&\text{Solution for $u$ (metavariable) found.}
% \end{aligned}
% \end{gather*}
% where the unification constraints and metavariable solutions are bound by a proper context.
% A unification problem in their system is described by $\Delta \Vdash \overline{K}$,
% where $\Delta$ is a collection of metavariables with their defined scopes.
% This approach clearly states the scroping of each unification problem,
% therefore rules out invalid instantiations to metavariables.
% 
% However, such ``duplicated contexts'' are not ideal for our formalization.
% Following DK's algorithm, an existential variable $\al$ defined in the context
% could be decomposed into a function type $\al[1] \to \al[2]$,
% so the declaration of $\al$ should be replaced by two declarations $\al[1], \al[2]$.
% Such operation causes all the context that contains $\al$ to change,
% which brings difficulty on the synchronization of multiple contexts,
% as variable declarations are not centralized.
% 
% We find a way of encoding multiple contexts by ``compressing'' them into a single worklist.
% As our type inference rules applied to a judgment,
% we typically keep the context or add variable declarations to the old context
% before analysing its sub-judgment(s).
% This enables us to write the worklist judgment in such a form:
% $$\Gm, \{\text{variable declarations}\}, \{\text{judgment}\},
% \{\text{variable declarations}\}, \{\text{judgment}\}, \ldots .$$
% Typically, we simplify the right-most judgment and push back smaller tasks,
% where the old ``context'' automatically gets inherited.
% When some rules solve or partially solve an existential variable,
% we could easily propagate the solution to all the judgments,
% and safely modify its declaration, such as remove from the worklist.
% New judgment(s) may also introduce new \emph{local} variables by declaring
% that variable right before the new judgment(s).
% After the new judgments are satisfied, these local variables are properly recycled.
% 
% % \jimmy{TODO explain the duplicated context with some examples (their judgment form)}
% % Some issues to point out: Duplicated contexts (rather than shared contexts), which make 
% % the formalization harder since requires ``synchronizing'' things ...
% 
%-------------------------------------------------------------------------------
\subsection{Single-Context Worklist Algorithm for Subtyping}
